---
title: "Reproducible documents"
author: "Fie Langmann"
format: html
editor: visual
---

To get code chunk: ctrl+alt+i

Create a new code chunk and then type setup right after the r. This section can be used to load packages and data

```{r setup}
#| message:false
#| warning:false
library(tidyverse)
library(NHANES)
nhanes_small <- read_csv(here::here("data/nhanes_small.csv"))
```
<!-- #| include:false will only run the code but not print anything in your document -->
Options added in the code chunk ony works for that specific code chunk and not for the others. If you want to add options for the whole document, they should be added in the YAML header instead as:
title: "Reproducible documents"
author: "Your Name"
format: html
execute:
  echo: false
  warning: false
  message: false
  
  
```{r}
nhanes_small
```
## Table of results

```{r mean-age-bmi-table}
# always insert an empty line after header before making new code chunk
nhanes_small %>%
  filter(!is.na(diabetes), !is.na(education)) %>%
  group_by(diabetes, education) %>%
  summarise(
    mean_age = mean(age, na.rm = TRUE),
    mean_bmi = mean(bmi, na.rm = TRUE)
  ) %>%
  ungroup() %>% # good practice to remember to ungroup things
  knitr::kable(caption = "Mean values of Age and BMI for education and diabetes status")
```
